---
title: "NEvent_postprocess"
author: "Nicole Deng"
output: html_document
date: "2023-04-29"
---
#Overview 
This R Markdown script serves as the first step in the data processing pipeline for the NEvent experiment. The script ingests raw output files produced by PsychoPy and conducts a range of data manipulations to ready the data for ensuing group effect studies and EEG analyses.

#Ouput
1. Individual datasets for encoding, temporal_memory, and source_memory (these are optional and the relevant code is currently commented out).
2. A merged dataset that includes all the data from the above three files, intended for integrated analyses.
3. A dataset containing behavioral results for individual participants.

```{r setup}
# Function to search for the CSV file of a specific subject and then load it
search_and_load_file <- function(subject_num) {
  
  # Create the directory path based on the subject number
  directory <- paste0("/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/sub-", sprintf("%03d", subject_num), "/beh/")
  
  # Use list.files to search for CSV files with the subject number in the specified directory
  files <- list.files(directory, pattern = paste0("^", subject_num, "_THESES_EVENT_FEDERICO_PSYCHOPY.*\\.csv$"), full.names = TRUE)
  
  # Check if the files exist
  if(length(files) == 0) {
    warning(sprintf("No CSV files found for subject number %d", subject_num))
    return(NULL)
  } else {
    # Return the file path (assuming there's only one match)
    file_path <- files[1]
    
    # Read the file into R
    data <- read.csv(file_path)
    
    # Return the data
    return(data)
  }
}

```

```{r}
# Test the function
subject_num <- 62
data <- search_and_load_file(subject_num)

if(!is.null(data)) {
  cat("The file is found, yay! CSV file for subject", subject_num, "has been loaded into 'data'.\n")

} else {
  cat("No CSV file found for subject", subject_num, "\n")
}
```

```{r}
library(dplyr)

#create the encoding file
 ### WHEN RUNNING EEG PARTICIPANTS ####
var.keep.encoding <- c("image_file", "category","boundary_type","tone_pitch",	"tone_source",	"list_num",	"event_num",	"serialpos_experiment",	"serialpos_list",	"serialpos_event","sound_1.started","stimulus_image.started","controller_encoding_resp.keys",	"controller_encoding_resp.rt","participant","date")

 ### WHEN RUNNING PURELY BEH PARTICIPANTS 
#var.keep.encoding <- c("image_file", "category","boundary_type","tone_pitch",	"tone_source",	"list_num",	"event_num",	"serialpos_experiment",	"serialpos_list",	"serialpos_event","controller_encoding_resp.keys",	"controller_encoding_resp.rt","participant","date")

encoding <- data [,var.keep.encoding ]
encoding <- encoding[complete.cases(encoding$tone_pitch),]
#encoding <- filter (encoding, list_num != 0)

### RANGES OF COLUMNS MIGHT NOT BE THE SAME FOR EVERY FILE

#create the temporal file
var.keep.temporal_memory <- c("image1_temp_order.started","controller_order_resp.keys",	"controller_order_resp.rt",	"image1_temp_distance.started","controller_distance_resp.keys",	"controller_distance_resp.rt","participant","date")
t1 <- data [,var.keep.temporal_memory]
t2 <- data [,11:33] 
temporal_memory <- cbind (t2, t1)
temporal_memory <- temporal_memory[complete.cases(temporal_memory$image1_list_num),]
temporal_memory<- filter (temporal_memory, image1_list_num != 0)

#create source file 
t3 <- data [,34:44] 
var.keep.source <- c("image_source.started",	"image_source.stopped",	"controller_source_resp.keys",	"controller_source_resp.rt","participant","date")
t4 <- data [,var.keep.source]
source_memory <- cbind (t3, t4)
source_memory <- source_memory[complete.cases(source_memory$source_image_tone_pitch),]
source_memory <- filter (source_memory, source_image_list_num != 0)

subjectnum = as.character(temporal_memory$participant[4])
```

## encoding
```{r}
encoding$TrueBoundary <- encoding$boundary_type
encoding$TrueBoundary <- ifelse(encoding$event_num == 0 & encoding$boundary_type == "B", "FT", encoding$boundary_type)
encoding$TrueBoundary [encoding$TrueBoundary=="2"] <-"NB"
encoding$TrueBoundary [encoding$TrueBoundary=="1"] <-"B"
encoding$TrueBoundary [encoding$TrueBoundary=="3"] <-"PreB"

encoding$BoundarySalience <- NA
for (i in 2:nrow(encoding)){
  encoding$BoundarySalience[i] <-encoding$tone_pitch[i]-encoding$tone_pitch[i-1]
}
```

## temporal_memory
### check the correctness 
```{r}
#var <-c("controller_order_resp.keys","left_image","right_image","correct_answer")
#temporal_memory_1 <- temporal_memory [,var] 
for (i in 1:nrow(temporal_memory)) {
  if (temporal_memory$controller_order_resp.keys[i] == 0) {
    temporal_memory$controller_order_resp.keys[i] <- temporal_memory$left_image[i]
  } else {
    temporal_memory$controller_order_resp.keys[i]<- temporal_memory$right_image[i]
  }
}
# append the col with correctness results back to main file
temporal_memory$correct <- ifelse(temporal_memory$controller_order_resp.keys == temporal_memory$correct_answer, 1, 0)
#temporal_memory_1$correct <- as.numeric(temporal_memory_1$correct)
#temporal_memory <- cbind (temporal_memory, temporal_memory_1$correct)
```

### Calculate pair Salience and distance rating
```{r}
#distance rating
if (subject_num%%2==0 || subject_num<43){ #if subject number is even OR smaller than 43 --> use the original code
  temporal_memory$controller_distance_resp.keys[temporal_memory$controller_distance_resp.keys == 2] <-1
} else { #subject number is odd
  temporal_memory$controller_distance_resp.keys[temporal_memory$controller_distance_resp.keys == 0] <-1
  temporal_memory$controller_distance_resp.keys[temporal_memory$controller_distance_resp.keys == 2] <-0
}

temporal_memory$PairSalience <- NA
for (i in 1:nrow(temporal_memory)){
  temporal_memory$PairSalience[i] <-temporal_memory$image2_tone_pitch[i]-temporal_memory$image1_tone_pitch[i]
}
```


## Source_memory
```{r}
source_memory$TrueBoundary <- ifelse(source_memory$source_image_event_num == 0 & source_memory$source_image_boundary_type == "B", "FT", source_memory$source_image_boundary_type)
source_memory$controller_source_resp.keys[source_memory$controller_source_resp.keys == 2] <-"R"
source_memory$controller_source_resp.keys[source_memory$controller_source_resp.keys == 0] <-"L"
source_memory$correct_source <- NA
source_memory$correct_source <- ifelse(source_memory$controller_source_resp.keys == source_memory$source_image_tone_source, 1, 0)
```

## export the datasets
```{r}

#filename1=paste("encoding_Sub",subjectnum,".csv")
#write.csv(encoding,filename1)

#filename2=paste("temporal_memory_Sub",subjectnum,".csv")
#write.csv(temporal_memory,filename2)

#filename3=paste("source_memory_Sub",subjectnum,".csv")
#write.csv(source_memory,filename3)

```

## merging the dataset
```{r}
#transform from wide to long
library(tidyr)
library(dplyr)
temporal_memory_long <- gather(temporal_memory,key="image_type",value="image_file",image1:image2)
temporal_memory_long <- select(temporal_memory_long, -(3:18))

#give encoding an id base of orders
#encoding <- encoding %>% mutate(id=1:n())
mergefile <- merge(encoding, temporal_memory_long, by="image_file",all=TRUE)
colnames (source_memory) [1]<- "image_file"
mergefile <- merge(mergefile, source_memory, by="image_file",all=TRUE)
mergefile <- arrange(mergefile,serialpos_experiment)

if (subject_num == 46) {
  mergefile <- mergefile[-(236:251), ] # Remove rows 236 to 251 due to EEG recording interruption
}

if (subject_num == 52) {
  mergefile <- mergefile[-(418:422), ] # Remove rows 236 to 251 due to EEG recording interruption
}

#merge file export
filename4=paste("mergefile_Sub-0",subjectnum,".csv",sep="")
pathname <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/mergefile"
full_path <- paste0(pathname,"/",filename4)
write.csv(mergefile, file = full_path)
```

##quick stats
```{r}

library(descr)
library(ggplot2)
print("distance rating")
tapply(temporal_memory$controller_distance_resp.keys,temporal_memory$pairtype,mean)
print("temporal order accuracy") 
tapply(temporal_memory$correct,temporal_memory$pairtype,mean)
print("source memory accuracy")
tapply(source_memory$correct_source,source_memory$TrueBoundary,mean)
print("encoding time")
encoding$controller_encoding_resp.rt <- as.numeric(encoding$controller_encoding_resp.rt)
tapply(encoding$controller_encoding_resp.rt, encoding$TrueBoundary, mean, na.rm=TRUE)

print('encoding reaction time vs. source accuracy')
tapply(mergefile$controller_encoding_resp.rt, factor(mergefile$correct_source), mean, na.rm=TRUE)

print('encoding reaction time vs. order accuracy')
tapply(mergefile$controller_encoding_resp.rt, factor(mergefile$correct), mean, na.rm=TRUE)
```
```{r}
# Create a new variable that combines pairtype and image2_list_num
temporal_memory$grouping_var <- interaction(temporal_memory$pairtype, temporal_memory$image2_list_num)

# Calculate the mean of controller_distance_resp.keys for each combination of pairtype and image2_list_num
distance_rating_by_list <- tapply(temporal_memory$controller_distance_resp.keys, temporal_memory$grouping_var, mean)

print (distance_rating_by_list)
# Load necessary libraries
library(ggplot2)

library(tidyr)


# Transform the output into a data frame
distance_df <- data.frame(mean_rating = unname(distance_rating_by_list), condition = names(distance_rating_by_list))

# Separate the condition into type ('B' or 'NB') and number
distance_df$type <- gsub("\\..*", "", distance_df$condition)
distance_df$number <- as.numeric(gsub(".*\\.", "", distance_df$condition))

# Convert the 'number' column to a factor with levels ordered
distance_df$number <- factor(distance_df$number, levels = sort(unique(distance_df$number)))

# Summarize mean ratings for 'B' and 'NB' separately for each number
summary_df <- distance_df %>%
  group_by(number, type) %>%
  summarize(mean_rating = mean(mean_rating), .groups = 'drop')

# Create a new data frame to determine where 'B' is greater than 'NB'
comparison_df <- summary_df %>%
  spread(key = type, value = mean_rating) %>%
  mutate(B_greater_NB = B > NB)

# Merge the comparison information back to the original data frame
distance_df <- distance_df %>%
  left_join(comparison_df %>% select(number, B_greater_NB), by = "number")

# Now create the bar plot with highlighting
ggplot(distance_df, aes(x = number, y = mean_rating, fill = interaction(B_greater_NB, type))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = c("FALSE.B" = "blue", "TRUE.B" = "red", "FALSE.NB" = "grey", "TRUE.NB" = "grey")) +
  labs(x = "Condition Number", y = "Mean Distance Rating", fill = "Condition Highlight") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x labels for clarity

# Display the plot
outputname=paste(subjectnum,".tiff",sep="")
path_to_save <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/distance_rating_by_list"
full_path1 <- paste0(path_to_save,"/",outputname)
tiff(full_path1, width = 900, height = 400, res=100)
last_plot()
dev.off()


```
# Data Aggregation and Export
```{r}
#
mean_distance_rating <- tapply(temporal_memory$controller_distance_resp.keys, temporal_memory$pairtype, mean)
mean_temporal_order_accuracy <- tapply(temporal_memory$correct, temporal_memory$pairtype, mean)
mean_source_memory_accuracy <- tapply(source_memory$correct_source, source_memory$TrueBoundary, mean)
mean_encoding_time <- tapply(encoding$controller_encoding_resp.rt, encoding$TrueBoundary, mean, na.rm=TRUE)


mean_reaction_time_vs_source <- tapply(mergefile$controller_encoding_resp.rt, factor(mergefile$correct_source), mean, na.rm=TRUE)
names(mean_reaction_time_vs_source)[names(mean_reaction_time_vs_source) == "0"] <- "Incorrect Source"
names(mean_reaction_time_vs_source)[names(mean_reaction_time_vs_source) == "1"] <- "Correct Source"

mean_reaction_time_vs_order <- tapply(mergefile$controller_encoding_resp.rt, factor(mergefile$correct), mean, na.rm=TRUE)
names(mean_reaction_time_vs_order)[names(mean_reaction_time_vs_order) == "0"] <- "Incorrect order"
names(mean_reaction_time_vs_order)[names(mean_reaction_time_vs_order) == "1"] <- "Correct order"

##
df_distance_rating <- data.frame(Category = names(mean_distance_rating),
                                 DistanceRating = mean_distance_rating)
df_temporal_order_accuracy <- data.frame(Category = names(mean_temporal_order_accuracy),
                                         TemporalOrderAccuracy = mean_temporal_order_accuracy)

df_source_memory_accuracy <- data.frame(Category = names(mean_source_memory_accuracy),
                                        SourceMemoryAccuracy = mean_source_memory_accuracy)
df_encoding_time <- data.frame(Category = names(mean_encoding_time),
                               EncodingTime = mean_encoding_time)
df_reaction_time_vs_source <- data.frame(Category = names(mean_reaction_time_vs_source),
                               Reaction_time_vs_source = mean_reaction_time_vs_source)

mean_reaction_time_vs_order <- data.frame(Category = names(mean_reaction_time_vs_order),
                               Reaction_time_vs_order = mean_reaction_time_vs_order)

# Merge these individual data frames into one wide-format data frame
merged_results <- Reduce(function(x, y) merge(x, y, by = "Category", all = TRUE),
                         list(df_distance_rating,
                              df_temporal_order_accuracy,
                              df_source_memory_accuracy,
                              df_encoding_time, df_reaction_time_vs_source, mean_reaction_time_vs_order))

# Convert merged data frame to a matrix
merged_matrix <- as.matrix(merged_results)

# Transpose the matrix
transposed_matrix <- t(merged_matrix)

# Convert the first row to column names
colnames(transposed_matrix) <- transposed_matrix[1, ]
transposed_matrix <- transposed_matrix[-1, , drop = FALSE]

# Convert the transposed matrix back to a data frame
final_results <- as.data.frame(transposed_matrix, stringsAsFactors = FALSE)

# Convert all numeric columns to numeric type (they will be read as characters initially)
numeric_cols <- colnames(final_results)
final_results[numeric_cols] <- lapply(final_results[numeric_cols], as.numeric)

#create ouput file
outputname=paste("results_Sub",subjectnum,".csv",sep="")
pathname1 <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/individual_results"
full_path1 <- paste0(pathname1,"/",outputname)
write.csv(final_results, file = full_path1)

```

