---
title: "Post_analysis"
author: "Nicole Deng"
date: "2023-09-16"
output: html_document
---
=======
Readme:
1. clean the individual result using exp_eeg_beh_processing_pipeline R code
2. On the current file, update the variable in the first block named "subject" to the current subject number. 
3. Run the rest of blocks
=======
Workflow
1. get input results files for all the existing participants
2. create four data frames [reaction_time, distance_rating, order_accuracy, source_accuracy]
3. align input files to the designed files 
4. connect the new data frames with existing statistical analyses/visualization codes
=======

#load packages
```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(scales)
```

# load the cleaned data
```{r}
subject <- c(36:61) 
df_combined <- data.frame()
for (i in subject) {
  filename <- paste("subject", as.character(i), sep="")
  subject_label <- paste("Sub", as.character(i), sep=" ")
  pathname <- paste ("/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/individual_results/results_Sub",as.character(i),".csv",sep="")
  temp_data <- read.csv(pathname)
  temp_data$Subject <- subject_label
  assign(filename, temp_data)
  df_combined <- rbind(df_combined, temp_data)
}
```

## create four data frames [reaction_time, distance_rating, order_accuracy, source_accuracy]
```{r }
colnames(df_combined)[1] <- "Metrics"

# Preparing distance rating data frame
distance_rating <- subset(df_combined, Metrics == "DistanceRating")
distance_rating <- distance_rating %>% select(-FT, -PreB)
distance_rating_long <- distance_rating %>% 
  pivot_longer(
    cols = c(B, NB), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )
write.csv(distance_rating, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/distance_rating.csv", row.names = FALSE)


# Preparing order_accuracy data frame
order_accuracy <- subset(df_combined, Metrics == "TemporalOrderAccuracy")
order_accuracy <- order_accuracy %>% select(-FT, -PreB)

order_accuracy_long <- order_accuracy %>% 
  pivot_longer(
    cols = c(B, NB), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )
write.csv(order_accuracy, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/order_accuracy.csv", row.names = FALSE)

# For the source_accuracy dataset
source_accuracy <- subset(df_combined, Metrics == "SourceMemoryAccuracy")
source_accuracy_long <- source_accuracy %>% 
  pivot_longer(
    cols = c(B, NB, FT, PreB), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )

write.csv(source_accuracy, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/source_accuracy.csv", row.names = FALSE)

# For the encoding_time dataset
encoding_time <- subset(df_combined, Metrics == "EncodingTime")
encoding_time_long <- encoding_time %>% 
  pivot_longer(
    cols = c(B, NB, FT, PreB), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )
write.csv(encoding_time, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/encoding_time.csv", row.names = FALSE)

# For the reaction_time_vs_source dataset
reaction_time_vs_source <- subset(df_combined, Metrics == "Reaction_time_vs_source")
reaction_time_vs_source_long <- reaction_time_vs_source %>% 
  pivot_longer(
    cols = c(Correct.Source,Incorrect.Source), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )


write.csv(encoding_time, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/reaction_time_vs_source", row.names = FALSE)

# For the reaction_time_vs_order dataset

reaction_time_vs_order <- subset(df_combined, Metrics == "Reaction_time_vs_order")
reaction_time_vs_order_long <- reaction_time_vs_order %>% 
  pivot_longer(
    cols = c(Correct.order,Incorrect.order), # Columns to be converted to long format
    names_to = "condition",    
    values_to = "value"      
  )


write.csv(encoding_time, "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/group_results/reaction_time_vs_order", row.names = FALSE)
```

# Temporal order distance ratings

### Participants as a block

#### t-test

```{r}
library(tidyr)
t.test(distance_rating$NB, distance_rating$B, paired=TRUE, alt="two.sided")
```

### Visualization

```{r, fig.width=4,fig.height=5}

ggplot(data=distance_rating_long) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=1) +
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.8) +
  geom_jitter(aes(x=condition, y=value), alpha=0.8, width=0, alpha=0.8) +
  #geom_text(aes(x=condition, y=value, label=Subject), vjust=-1, size=3) +
  scale_fill_manual(values=c("#306690", "#5E6469")) +
  ggtitle("Distance rating")+
  ylim(0, 1)

path_to_save <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/distance_rating.jpg"
jpeg(path_to_save, width = 1200, height = 1200, res=300)
last_plot()
dev.off()

##
#ggpaired(distance_rating_long, x="condition", y="value",
 #        color="condition", line.color="grey",
#         line.size=0.4)+
#  ylim(0, 1)+
#  stat_compare_means(method = "t.test",paired = TRUE) + 
#  geom_jitter(aes(color = condition), position = position_jitterdodge(jitter.width = 0), alpha = 0.8) + 
#  scale_color_manual(values=c("#5E6469", "#5E6469"), labels = c("Boundary Pairs", "Non-Boundary Pairs")) +
#  geom_boxplot(aes(fill = condition), width = 0.5, alpha = 0.8) +
#  scale_fill_manual(values=c("#306690", "#5E6469"), labels = c("Boundary Pairs", "Non-Boundary Pairs")) +
#  guides(color=guide_legend(override.aes=list(size=1)),
#         fill=guide_legend(override.aes=list(size=1)))+
#  labs(y = "distance_rating")



```

# Average encoding reaction time

### ANOVA +block

```{r}
#add the block
model4= aov(value~condition+as.factor(Subject), data=encoding_time_long)
summary(model4)
layout(matrix(c(1,2,3,4), 2,2))
plot(model4)
```

### Visualization

```{r, fig.width=7,fig.height=7}
encoding_time_long$condition <- factor(encoding_time_long$condition, levels = c("B", "NB", "FT", "PreB"))
ggplot(data=encoding_time_long) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=0.7) + # Connect dots of same participant
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.7) +
  geom_jitter(aes(x=condition, y=value), alpha=0.6, width=0) +
  #geom_text(aes(x=condition, y=value, label=Subject), vjust=-1, size=2) + # Label each participant
  ggtitle("Encoding time")+
  scale_fill_manual(values=c("#306690",  "#5E6469", "skyblue1","white")) +
  ylim(0, 2.2)

#save the plots to tiff format
path_to_save <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/encoding_time.tiff"
tiff(path_to_save, width = 500, height = 400,res=90)
last_plot()
dev.off()
```



# Order accuracy

### Participants as a block

#### t-test

```{r}
t.test(order_accuracy$NB, order_accuracy$B, paired=TRUE, alt="two.sided")
```
### Visualization
```{r, fig.width=4,fig.height=4}
library(ggpubr)

ggpaired(order_accuracy_long, x="condition", y="value",
         color="condition", line.color="grey",
         line.size=0.4) +
  ylim(0, 1) +
  stat_compare_means(method = "t.test", paired = TRUE) + 
  geom_jitter(aes(color = condition), position = position_jitterdodge(jitter.width = 0), alpha = 1) + 
  scale_color_manual(values=c("#5E6469", "#5E6469"), labels = c("Boundary Pairs", "Non-Boundary Pairs")) +
  geom_boxplot(aes(fill = condition), width = 0.5, alpha = 0.7) +
  scale_fill_manual(values=c("#306690", "#5E6469"), labels = c("Boundary Pairs", "Non-Boundary Pairs")) +
  geom_text(aes(x=condition, y=value, label=Subject), vjust=-1, size=3) + # Added this line for labeling+
  labs(y = "Order accuracy")

ggplot(data=order_accuracy_long) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=0.9) +
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.5) +
  geom_jitter(aes(x=condition, y=value), alpha=0.7, width=0) +
  #geom_text(aes(x=condition, y=value, label=Subject), hjust=-0.1, vjust=-1, size=3) +
  scale_fill_manual(values=c("#306690", "#5E6469")) +
  ggtitle("Order accuracy")+
  ylim(0, 1)

path_to_save <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/order_accuracy.jpg"
jpeg(path_to_save, width = 1200, height = 1200, res=300)
last_plot()
dev.off()


```
# Source accuracy

### Participants as a block

#### Anova

```{r}
model5= aov(value~condition+factor(Subject),
            data=source_accuracy_long)
summary(model5)

t.test(source_accuracy$NB, source_accuracy$B, paired=TRUE, alt="two.sided")
```

```{r}
layout(matrix(c(1,2,3,4), 2,2))
plot(model5)
```
Graph all
```{r,fig.width=6,fig.height=6}
source_accuracy_long$condition <- factor(source_accuracy_long$condition, levels = c("B", "NB", "FT", "PreB"))
ggplot(data=source_accuracy_long)+
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.5)+
  geom_jitter(aes(x=condition, y=value),alpha = 0.4, width=0.2, size=1)+
  scale_fill_manual(values=c("#306690", "#5E6469","skyblue1","white"))+
  #geom_text(aes(x=condition, y=value, label=Subject), hjust=-0.1, vjust=-1, size=2) +
  ggtitle("Source_accuracy")+
  ylim(0, 1)

path_to_save <- "/Volumes/ezzyatlab/experiments/NEvent/exp_eeg_v1/eeg_beh_results/Source_accuracy.jpg"
jpeg(path_to_save, width = 1200, height = 1200, res=300)
last_plot()
dev.off()
```
Graph only B and NB items

```{r, fig.width=4,fig.height=5}
filtered_data <- source_accuracy_long %>% filter(condition %in% c('B', 'NB'))

ggplot(data=filtered_data) +
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.4) +
  geom_jitter(aes(x=condition, y=value),alpha = 0.9, width=0) +
  scale_fill_manual(values=c("#306690", "#5E6469","skyblue1","white")) +
  geom_text(aes(x=condition, y=value, label=Subject), hjust=-0.1, vjust=-1, size=3) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=0.9) +
  ylim(0, 1)
```
Correlation: beh_Source_memory vs. beh_order accuracy
```{r }



```





Graph reaction_time_vs_source

```{r, fig.width=2,fig.height=4}
#filtered_data <- source_accuracy_long %>% filter(condition %in% c('B', 'NB'))

ggplot(data=reaction_time_vs_source_long) +
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.4) +
  geom_jitter(aes(x=condition, y=value),alpha = 0.9, width=0) +
  scale_fill_manual(values=c("#306690", "#5E6469")) +
  geom_text(aes(x=condition, y=value, label=Subject), hjust=-0.1, vjust=-0.5, size=3) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=0.9)+
  ggtitle("Reaction time vs source accuracy")
```
Graph reaction_time_vs_order

```{r, fig.width=2,fig.height=2}
#filtered_data <- source_accuracy_long %>% filter(condition %in% c('B', 'NB'))

ggplot(data=reaction_time_vs_order_long) +
  geom_boxplot(aes(x=condition, y=value, fill=condition), alpha=0.4) +
  geom_jitter(aes(x=condition, y=value),alpha = 0.9, width=0) +
  scale_fill_manual(values=c("#306690", "#5E6469")) +
  geom_text(aes(x=condition, y=value, label=Subject), hjust=-0.1, vjust=-0.5, size=3) +
  geom_line(aes(x=condition, y=value, group=Subject), alpha=0.9)+
  ggtitle("Reaction time vs order accuracy")+
  ylim(0,2)
```
